Three 38 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia

### 2022-03-16 papers/ms.tex
update ms: Gaia selection procedure

@@

The discovery of young transiting planets is a major frontier in
current exoplanet research.  The reason is that the properties of
these planets can help constrain the timescales for processes
including hot Jupiter migration (CITE), the cooling and contraction of
mini-Neptunes (CITE Rizzuto), the early evolution of close-in
multiplanet systems (CITE David), and whether and how mass-loss
explains the radius valley (e.g., CITE, CITE, Lee+21 on theory; CITE
CITE CITE on observations).

%FIXME WORDING OF THIS PARAGRAPH NEEDS WORK.  ALSO, WHERE DO YOU PUT
%THE CONTEXT RE CLUSTERS?
Understanding the formation and evolution of stellar clusters is
similarly an emerging frontier in analyses of star formation and
cluster evolution.  In particular, it is understood that most embedded
clusters disperse within a few tens of millions of years after forming
their first stars (CITE).  The discovery of tidal tails (CITE CITE)...
streams (CITE Meingast18, Curtis19), etc...  Connect to moving groups
(CITE Gagne 21).  What are the typical densities for star formation?
How does gas expulsion and the SFE affect cluster morphology?  {\bf
FIXME}

Assuming a constant star formation rate in the Galaxy, $\approx$1\% of
stars are $\lesssim$10$^8$ years old.  While $\approx$5{,}000
transiting exoplanets are known (CITE), 11 currently meet this age
cut.  The discrepancy between the $\approx$50 expected and $\approx$10
known likely has a contribution from selection effects in planetary
detection, which include photometric and spectroscopic starspot
modulation, and the rarity of young stars.

While detecting planets around young and active stars is challenging,
so is measuring stellar ages (see CITE Soderblom for a review).  It
should therefore not be surprising that advances along either
dimension can yield advances in the census of young transiting
planets.

The prime Kepler mission (CITE) found most of the currently known
transiting exoplanets, and it was conducted before the positions,
velocities, and stellar brightnesses measured by Gaia revolutionized
our understanding of stellar clusters (CITE, CITE, CITE).  This
revolution is slated to continue with the upcoming third Gaia data
release (DR3 CITE).  Nonetheless, it seems sensible to revisit the
Kepler field, given our new knowlege of the stellar ages.

In this work, we expand on a previous study of a 38 million year (Myr)
old Neptune-sized planet in the Kepler field (Kepler-1627Ab; CITE).
The age of this planet was derived based on its host star's membership
in the $\delta$\ Lyr\ cluster.  Our analysis of this cluster focused
on the immediate kinematic vicinity of Kepler-1627A in order to
reliably confirm the age of the planet.  However it also became clear
that the $\delta$\ Lyr\ cluster is spatially and kinematically close
to a much larger group of similarly aged stars.  This group, which is
at a distance of roughly 300 parsecs, appears to span Cepheus
($l=100^\circ$) to Hercules ($l=40^\circ$), at galactic latitudes,
$b$, between roughly 0 and 20 degrees.  We therefore refer to it as
the Cep-Her complex.  Viewed top-down in the plane of the Galaxy, the
complex spans roughly 250 parsecs in the direction toward the galactic
center, and 150 parsecs in the perpendicular direction of galactic
rotation.  A detailed kinematic analysis and exploration is currently
being prepared by R.~Kerr and collaborators.

The subject of this study is the intersection of the Cep-Her complex
with the Kepler field, and an analysis of the consequent transiting
planets and planet candidates.  Cross-matching the full set of
candidate Cep-Her members against all known Kepler Objects of Interest
(CITE) yielded XX candidate matches.  Given the purported XX Myr age
of the group, a visual inspection of the light curves quickly enabled
a decision about whether the matches could be valid, based on the
presence or non-presence of stellar rotation signals of the
appropriate periods and amplitudes (CITE REBULL).  Four possible
matches remained from this exercise: Kepler-1627, Kepler-1643,
KOI-7368, and KOI-7913.  Given our previous analysis of Kepler-1627,
we will focus for the most part on the latter three.

@@

\subsection{Selecting Cluster Members}
\label{sec:kinematicselection}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Galactic positions  and tangential velocities of stars in the
    complex.}  {\bf This is a placeholder before we get real kinematics.}
    Sub-clusters include the $\delta$ Lyr cluster, RSG-5, and the
    worryingly diffuse ``CH-2''.
		\label{fig:XYZvtang}
	}
\end{figure*}

Our focus is on a region of the Galaxy roughly 200 to 500 parsecs from
the Sun, above the galactic plane, and spanning galactic longitudes of
roughly $40^\circ$ to $100^\circ$ degrees.  A direct query of the Gaia
archive imposing these constraints yields $\approx3.4\times10^7$
sources.  Imposing the cuts suggested by \citet[][Appendix
B]{GaiaCollaboration2018} on the photometric signal to noise ratio,
the number of visibility periods used, the astrometric $\chi^2$ of the
single-source solution, the $G_{\rm BP}-G_{\rm RP}$ color excess
factor and the parallax precision ($\varpi/\sigma_\varpi>5$) yields
$\approx5.0\times10^6$ sources.

Imposing a thin slice in color-absolute magnitude space around the
pre-main-sequence of the $\delta$ Lyr cluster's previously established
$\approx$40 Myr locus, and imposing kinematics cut for $-8 < v_b /
{\rm km\,s^{-1}} < 0$ and $-15 < v_{l'} / {\rm km\,s^{-1}} < 15$,
where $l'=l\cos b$, yields $\approx 5.2\times10^3$ sources.  These
objects are all in the range BP-RP 1.5 to 3.2, so roughly K5V to
M4.5V, or $\approx$0.2 to 0.7 solar masses.  Much more manageable, but
according to IMF arguments, missing at least a factor of two in mass
of any clusters.

In XYZ, and in tangential velocity space, these five thousand sources
are visually clustered in between a few to a dozen groups, depending
on the characteristic size and velocity scales of interest.  The first
suggestion that these objects could all be part of a monolithic
structure, to our knowledge, came from the manual grouping that was
performed between them by \citet{KounkelCovey2019} when defining the
structure they dubbed ``Theia~73''.  Different subsets of these stars
have however previously been noticed and studied.  They include the
$\delta$ Lyr cluster, also known as Stephenson-1
\citep{stephenson_possible_1959}, which comprises a major grouping
around $(l,b)=(68^\circ,15^\circ)$, with typical velocities $(v_{l'},
v_b)=(-4.5{\rm \kms},-4{\rm \kms})$.  Another individual component
that has previously been recognized is RSG-5 \citep{roser_nine_2016}.
This group is at roughly $(l,b)=(83^\circ,6^\circ)$, with velocities
$(v_{l'}, v_b)=(+5.5{\rm \kms},-3.5{\rm \kms})$.  Most of the other
subclusters, including CH-2, and those in Cep-Cyg and Cerberus are too
small or dispersed to have previously been analyzed in great detail.

>>>>

\newcommand{\npms}{1097} % 20220311_Kerr_SPYGLASS205_Members_All.csv

@@

The discovery and characterization of transiting planets younger than
a billion years is a major frontier in current exoplanet research.
The reason is that the properties of young planets provide benchmarks
for studies of planetary evolution.  For instance, there are the
questions of when hot Jupiters arrive on their close-in orbits
\citep{dawson_johnson_2018}, how the sizes of planets with massive
gaseous envelopes evolve \citep{rizzuto_tess_2020}, when and if
close-in multiplanet systems fall out of resonance
\citep{arevalo_stability_2022,goldberg_architectures_2022},
and whether and how mass-loss explains the radius valley
\citep{lopez_how_2012,Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018,lee_primordial_2021}.

The discovery of a young planet requires two claims to be fulfilled:
the planet must exist, and its age must be secured.  Spaced-based
photometry from K2 and TESS has yielded a number of young planets for
which the planetary evidence comes from transits, and the age evidence
is based on either cluster membership \citep[{\it
e.g.},]{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
or else on correlates of youth such as stellar rotation, photospheric
lithium abundances, x-ray activity, or emission line strength
\citep[{\it e.g.},]{zhou_2021_tois,hedges_toi-2076_2021}.

In this work, we leverage recent analyses of the Gaia data, which have
greatly expanded our knowledge of stellar group memberships
\citep[{\it e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.
Broadly speaking these analyses cluster on the stellar positions and
on-sky velocities measured by Gaia, with varying degrees of filtering
and supervision.  One important result is the identification of
diffuse streams and tidal tails comparable in stellar mass to the
previously known cores of many open clusters
\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  These
spatially diffuse groups typically have velocity dispersions of
$\sim$1\kms, and can be verified to be ``co-eval'' through analyses of
color-absolute magnitude diagrams, stellar rotation periods
\citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
\citep{hawkins_2020}.  While there are many implications for our
understanding of star formation and cluster evolution \citep[{\it
e.g.}][]{dinnbier_tidal_2020}, a more immediate consequence is that we
now know the ages of many more stars, including previously known
planet hosts.

%Assuming a constant star formation rate in the Galaxy, $\approx$1\% of
%stars are $\lesssim$10$^8$ years old.  While $\approx$5{,}000
%transiting exoplanets are known (CITE), 11 currently meet this age
%cut.  The discrepancy between the $\approx$50 expected and $\approx$10
%known likely has a contribution from selection effects in planetary
%detection, which include photometric and spectroscopic starspot
%modulation, and the rarity of young stars.

% While detecting planets around young and active stars is challenging,
% so is measuring stellar ages (see CITE Soderblom for a review).  It
% should therefore not be surprising that advances along either
% dimension can yield advances in the census of young transiting
% planets.

The prime Kepler mission \citep{borucki_kepler_2010} found most of the
currently known transiting exoplanets, and it was conducted before
Gaia.  It therefore seems sensible to revisit the Kepler field, given
our new knowlege of the stellar ages.

In this work, we expand on our previous study of a $38^{+7}_{-6}$
million year old Neptune-sized planet in the Kepler field (Kepler-1627
Ab; \citealt{bouma_kep1627_2022}).  The age of this planet was derived
based on its host star's membership in the $\delta$\ Lyr\ cluster.
Our analysis of the cluster focused on the immediate spatial and
kinematic vicinity of Kepler-1627~A in order to reliably confirm the
age of the planet.  However it became clear that the $\delta$\ Lyr\
cluster seems to also be part of a much larger group of similarly aged
stars.  This group, which is at a distance of $\approx$300 parsecs
from the Sun, appears to span Cepheus to Hercules (galactic
longitudes, $l$, between 40$^\circ$ and 100$^\circ$), with galactic
latitudes spanning roughly 0 and 20 degrees.  We therefore refer to it
as the Cep-Her complex.  It exhibits significant sub-structure over
its $\approx$250 parsec length, and a detailed analysis that delves
into the memberships, kinematics, and possible origin of the group is
currently being prepared by R.~Kerr and collaborators.

Here, our focus is on the intersection of the Cep-Her complex with the
Kepler field.  Cross-matching the full set of candidate Cep-Her
members against known Kepler Objects of Interest
\citep{thompson_planetary_2018} yielded four candidate cluster
members: Kepler-1627, Kepler-1643, KOI-7368, and KOI-7913.  Given our
previous analysis of Kepler-1627, we will mostly focus on the latter
three.  After analyzing the relevant properties of Cep-Her
(Section~\ref{sec:cluster}), we discuss the stellar properties
(Section~\ref{sec:stars}) and validate the planetary nature of each
system using a combination of the Kepler photometry and
high-resolution imaging (Section~\ref{sec:planets}).  We conclude with
a discussion of implications for the size-evolution of close-in
mini-Neptunes (Section~\ref{sec:disc}).

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
  {\bf Positions and velocities of candidate members of the Cep-Her
  complex.}
  {\it Top row}: On-sky positions in galactic coordinates.  Black
  points are stars for which group membership is more likely than for
  gray points (see Section~\ref{subsec:members}).
  Kepler-1627 is in the outskirts of the $\delta$ Lyr cluster
  \citep{bouma_kep1627_2022}, which is centered at $\{ l, b\} \approx
  \{ 66^\circ, 12^\circ\}$.
  {\it Middle row}: Galactic positions.  The Sun is at $\{X, Y, Z\} =
  \{0, 0, 20.8\}$ pc.
  {\it Bottom row}: Galactic tangential velocities (left) and
  longitudinal tangential velocity versus galactic latitude (right).
  The gray band in the lower-right shows the $\pm$1-$\sigma$
  projection of the Solar velocity with respect to the local standard
  of rest.
  There is a strong spatial and kinematic overlap between Kepler-1643
  and RSG-5 (magenta points; $\{ l, b\} \approx \{ 82^\circ,
  6^\circ\}$; $\{ X, Y, Z\} \approx \{-8070, 330, 50\} $ pc).  The
  local population of candidate young stars around KOI-7368 and
  KOI-7913 is more diffuse -- we call this region ``CH-2'' (lime green
  points).  The green and purple points are manually selected from the
  initial candidate population based on galactic position and
  tangential velocity.
	\label{fig:XYZvtang}
	}
\end{figure*}

\subsection{Previous Related Work}

Our focus is on a region of the Galaxy roughly 200 to 500 parsecs from
the Sun, above the galactic plane, and spanning galactic longitudes of
roughly $40^\circ$ to $100^\circ$ degrees.  Two of the richest
clusters in this region are the $\delta$~Lyra cluster
\citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
Both of these clusters were known before Gaia, and have reported ages
between $\approx$30 and $\approx$60 million years.  Early
empirical evidence that these two clusters might be part of a broader
population was apparent in the Gaia-based photometric analysis of
pre-main-sequence stars by \citet[][see their Figures~11 and~13]{Zari2018}.
Further kinematic connections and complexity were highlighted by
\citet{KounkelCovey2019}, when defining the structures they dubbed
``Theia~73'' and ``Theia~96''\footnote{A useful visualization is
available online at \url{http://mkounkel.com/mw3d/mw2d.html} (accessed
15 March 2022)}.  This connection by \citet{KounkelCovey2019} was made
as part of an unsupervised clustering analysis of the Gaia DR2
positions and tangential velocities with a subsequent manual
``stitching'' step, and generally supports the idea that there is an
overdensity of $30$--$60\,{\rm Myr}$ old stars in this region of the
Galaxy.  \citet{Kerr2021}, in a volume-limited analysis of the Gaia
DR2 point-source catalog out to one third of a kiloparsec, identified
three of the nearest sub-populations, dubbed ``Cepheus-Cygnus'',
``Lyra'', and ``Cerberus''.  \citet{Kerr2021} reported ages for each
of these subgroups between 30 and 35 million years.

\subsection{Member Selection}
\label{subsec:members}

{\bf TODO RONAN: PLEASE VET ENTIRE SECTION, AND UPDATE WHERE
APPROPRIATE!}.

The possibility that the $\delta$~Lyr cluster, RSG-5, and a few nearby
sub-populations in this quadrant of the sky share a common origin has
yet to be fully substantiated.  To further explore the population of
$10$--$100\,{\rm Myr}$ stars in this region, we select candidate
members through four steps, the first three being identical to those
described in Section~3 of \citet{Kerr2021}.  We briefly summarize them
here.

The first step is to select stars that are photometrically distinct
from the field star population based on Gaia EDR3 magnitudes $\{G,
G_{\rm RP}, G_{\rm BP}\}$, parallaxes and auxiliary reddening
estimates \citep{lallement_gaia-2mass_2019}.  This step yielded \npms\
stars, which are either  pre-main-sequence K and M dwarfs due to their
long contraction timescales, or massive stars near the zero-age main
sequence due to their rapid evolutionary timescales.  The second step
is to then perform in an unsupervised HDBScan clustering on this
photometrically-young population
\citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The specific
parameters we use in this clustering analysis are $\{ X, Y, Z, c v_b,
c v_{l^*} \} $, where $c$ is the size-velocity corrective factor,
which is taken as $c=6\,{\rm pc / km\,s}^{-1}$ to ensure that the
spatial and velocity scales have identical standard deviations.  As
input parameters, we set the minimum $\epsilon$ threshold past which
clusters cannot be fragmented as $25$ in units of parsecs, and $c$
\kms\ in units of velocity.  The minimum cluster size $N$ is set to
10, as is $k$, the parameter used to define the ``core distance''
density metric.  This unsupervised clustering in our case
yielded 8 distinct groups.
These groups are then used as the ``seed'' populations for the third
step, which is to search for objects that are spatially and
kinematically consistent with each of the groups, {\bf WORDING NEEDS
WORK FOR CLARITY at least as close to the 10$^{\rm th}$ nearest HDBSCAN-identified
member in space-velocity coordinates as the most peripheral
HDBSCAN-identified member}.
The latter step yields F, G, and K-type stars that are spatially and
kinematically close to the photometrically-young stars, but which
cannot be identified as young based on their positions in color versus
absolute magnitude.

The outcome of this analysis is shown in Figure~\ref{fig:XYZvtang}.
We computed positions in these plots assuming the \texttt{astropy
v4.0} coordinate standard \citep{astropy_2018}, which assumes that the
Sun is $8122$ pc from the galactic center.  It also assumes
the solar velocity with respect to the local standard of rest from \citet{schonrich_local_2010}.
The previously known $\delta$~Lyr\ cluster ($l,b=68^\circ,15^\circ$;
$v_{l'}, v_b=-4.5{\rm kms},-4{\rm kms}$) is visible, as is
RSG-5 ($l,b=83^\circ,6^\circ$, $v_{l'}, v_b=+5.5{\rm kms},-3.5{\rm
kms}$).  Most of the other subclusters, including CH-2, and those in
Cep-Cyg ($l,b=90^\circ,7^\circ$) and Cerberus
($l,b=48^\circ,18^\circ$) are too small or dispersed to have
previously been analyzed in great detail.

To enable a selection cut that filters out field-star contaminants, we
also compute a weight metric, defined such that the group member
with the smallest core distance has a weight of 1, the group member
with the greatest core distance has a weight of 0, and weights for the
other group members are linearly spaced between these two extremes.
For Figure~\ref{fig:XYZvtang}, we show 26{,}960 objects with weight
exceeding 0.02 as gray points, and overplot 7{,}560 objects with
weights exceeding 0.10 as black points.

% FIXME FIXME ANYTHING ELSE TO SAY?
{\bf TODO: describe final step: manual selection of RSG-5 and CH-2 populations}

%Querying the Gaia archive
%and imposing the cuts suggested by \citet[][Appendix
%B]{GaiaCollaboration2018} on the photometric signal to noise ratio,
%the number of visibility periods used, the astrometric $\chi^2$ of the
%single-source solution, the $G_{\rm BP}-G_{\rm RP}$ color excess
%factor and the parallax precision ($\varpi/\sigma_\varpi>5$) yields
%$\approx5.0\times10^6$ sources.
%
%Imposing a thin slice in color-absolute magnitude space around the
%pre-main-sequence of the $\delta$ Lyr cluster's previously established
%$\approx$40 Myr locus, and imposing kinematics cut for $-8 < v_b /
%{\rm km\,s^{-1}} < 0$ and $-15 < v_{l'} / {\rm km\,s^{-1}} < 15$,
%where $l'=l\cos b$, yields $\approx 5.2\times10^3$ sources.  These
%objects are all in the range BP-RP 1.5 to 3.2, so roughly K5V to
%M4.5V, or $\approx$0.2 to 0.7 solar masses.  Much more manageable, but
%according to IMF arguments, missing at least a factor of two in mass
%of any clusters.
%
%In XYZ, and in tangential velocity space, these five thousand sources
%are visually clustered in between a few to a dozen groups, depending
%on the characteristic size and velocity scales of interest.
